Product Price Analysis Using Observational Data

### 2024-11-14 paper/paper.qmd
Complete paper.

---
title: "Product Price Analysis Using Observational Data"
subtitle: "Analysis of Price Trends, Minimal Correlation with Unit Pricing, and Potential Data Biases"
author: 
  - Yizhe Chen
  - Another author
thanks: "Code and data are available at: [https://github.com/YizheChenUT/Product_Prices.git](https://github.com/YizheChenUT/Product_Prices.git)."
date: today
date-format: long
abstract: "This report provides an analysis of observational product price data, focusing on current price statistics, price changes, and the weak correlation observed between current price and unit price. The findings indicate that there is minimal linear relationship between these two variables, suggesting that other factors may play a more substantial role in determining product prices. In the discussion, we address the implications of this weak correlation, the impact of missing data on price change calculations, and potential sources of bias that could influence the analysis."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

library(tidyverse)

@@

This report presents an analysis of product price data obtained from the hammer-2-processed.sqlite database. The primary objective of this analysis is to explore price trends, calculate price changes over time, and examine the potential correlation between current price and price per unit. The results show a very weak negative correlation between these two variables, indicating that changes in price per unit have little to no linear relationship with current price. This observational data, sourced from multiple vendors, provides insights into pricing behavior but also presents challenges, such as missing data and potential sources of bias, which could impact the reliability of the findings.

@@

The dataset used in this analysis includes key columns such as current price, old price, and price per unit. These columns allow us to calculate basic statistics on product pricing, examine price changes over time, and investigate any potential relationship between unit pricing and overall price. The data was collected from various vendors, providing a range of pricing practices and helping to create a comprehensive view of pricing strategies.

@@

- `current_price`: The most recent recorded price of the product.

- `old_price`: The previous price recorded for each product, which is used to calculate the percentage change in price over time.

- `price_per_unit`: The unit price of each product, which could theoretically be associated with the overall current price.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-basic-stats
#| tbl-cap: "Basic statistics of current and old prices"
# Read data
basic_stats <- read.csv(here::here("data/02-analysis_data/result_analysis1.csv"))

# Display of basic statistical data tables
knitr::kable(basic_stats)

@@

@fig-price-change shows the percentage change in price for each product where `old price` is available.

@@

# Read data on price change in per cent
price_change <- read.csv(here::here("data/02-analysis_data/result_analysis2.csv"))

>>>>

---
title: "Product Price Analysis Using Observational Data"
subtitle: "Analysis of Price Trends, Minimal Correlation with Unit Pricing, and Potential Data Biases"
author: 
  - Yizhe Chen
  - Bo Tang
  - Qizhou Xie
thanks: "Code and data are available at: [https://github.com/YizheChenUT/Product_Prices.git](https://github.com/YizheChenUT/Product_Prices.git)."
date: today
date-format: long
abstract: "This report provides an analysis of observational product price data, focusing on current price statistics, price changes, and the weak correlation observed between current price and unit price. The findings indicate that there is minimal linear relationship between these two variables, suggesting that other factors may play a more substantial role in determining product prices. In the discussion, we address the implications of this weak correlation, the impact of missing data on price change calculations, and potential sources of bias that could influence the analysis."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

library(tidyverse)
library(kableExtra)

@@

This report presents an analysis of product price data obtained from the Project Hammer – Jacob Filipp [@filipp]. We use the Structured Query Language (SQL) [@sql] to process and analyze and apply the statistical programming language R [@citeR] to make tables and graphs. The primary objective of this analysis is to explore price trends, calculate price changes over time, and examine the potential correlation between current price and price per unit. The results show a very weak negative correlation between these two variables, indicating that changes in price per unit have little to no linear relationship with current price. This observational data, sourced from multiple vendors, provides insights into pricing behavior but also presents challenges, such as missing data and potential sources of bias, which could impact the reliability of the findings.

@@

The dataset used in this analysis from the Project Hammer – Jacob Filipp [@filipp] includes key columns such as current price, old price, and price per unit. These columns allow us to calculate basic statistics on product pricing, examine price changes over time, and investigate any potential relationship between unit pricing and overall price. The data was collected from various vendors, providing a range of pricing practices and helping to create a comprehensive view of pricing strategies.

@@

- current price(current_price): The most recent recorded price of the product.

- old price(old_price): The previous price recorded for each product, which is used to calculate the percentage change in price over time.

- price per unit(price_per_unit): The unit price of each product, which could theoretically be associated with the overall current price.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-basic-stats
#| tbl-cap: "Basic statistics of current and old prices"
# Read data
basic_stats <- read.csv(here::here("data/01-analysis_data/result_analysis1.csv"))

# Shortening of column headings
colnames(basic_stats) <- c("Avg Curr Price", "Max Curr Price", "Min Curr Price",
                           "Avg Old Price", "Max Old Price", "Min Old Price",
                           "Avg Price Diff", "Max Price Diff", "Min Price Diff")

# Display of basic statistical data tables
basic_stats %>%
  kable("latex", booktabs = TRUE, linesep = "") %>%
  kable_styling(font_size = 4.5, full_width = FALSE, position = "center")

@@

@fig-price-change shows the percentage change in price for each product where `old price` is available. We use the tidyverse package [@tidyverse] to make the figure.

@@

# Read data on price change in per cent
price_change <- read.csv(here::here("data/01-analysis_data/result_analysis2.csv"))
